Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-10-31 manuscript.tex
edits to manuscript that are mentioned in responses

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute \change{a small, a big, or a huge strawberry } to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average \change{smaller} strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options -- ``TRUE'' and ``FALSE' -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly ``FALSE' responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypotheses}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). While some studies have used
  \enquote{false}--\enquote{true} \cite{bott2004, doran2012}, others have argued that judging truth
  may lead to meta-linguistic reasoning in participants that could
  distort judgments \change{\cite{XXX} XXX FIX}.} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

@@

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) with a weak link resulted in greater
implicature rates. Finally, there was a three-way interaction between
link, trial type, and number of response options (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). One interpretation of this
interaction is that the difference between the Weak and Strong link on
scalar trials in the quinary task was smaller than on exhaustive trials,
though we believe this is not too interesting, given that the binary
reference level implicature estimate was lower for exhaustive trials in
the first place. Crucially, both number of response options and link
affect the inferred implicature rate.

@@

>>>>

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011); \change{Sikos et al (this issue)}). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute \change{a small, a big, or a huge strawberry } to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average \change{smaller} strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options---``TRUE'' and ``FALSE'---to interpretations---literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly ``FALSE' responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypotheses}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
\change{assumes} two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). \change{Studies vary in the labeling of scale points.}} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

@@

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) with a weak link resulted in greater
implicature rates \change{than with a strong link}. Finally, there was a three-way interaction between
link, trial type, and number of response options (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). One interpretation of this
interaction is that the difference between the Weak and Strong link on
scalar trials in the quinary task was smaller than on exhaustive trials,
though we believe this is not too interesting, given that the binary
reference level implicature estimate was lower for exhaustive trials in
the first place. Crucially, both number of response options and link
affect the inferred implicature rate.

@@

\change{\hypertarget{ref-Sikos2019}{}
Sikos, L., Kim, M., \& Grodner, D.J. (in press). Social context modulates tolerance for pragmatic violations in binary but not graded judgments. \emph{Frontiers in Psychology}.}
